---
title: "Prediction Assignment Writeup"
author: "Fred Bosman"
date: "2 april 2016"
output: html_document
---

##Summary

In this report we will examine data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. We will try to predict which of the ways the was used.

##Load and clean data

First we set the working directory and load the necessary libraries.

```{r libaries}
setwd("C:/Fred/_werk/Coursera/PracticalMachineLearning_PredictionAssignment")
library(caret)
library(randomForest)
```

Next we will download the files and load them into R. Because the CSV-file containing the training data is large, we will first check if it's already downloaded. 

```{r load_files}
# check if file is already downloaded
if (!file.exists("pml-training.csv")) {
  trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

  download.file(trainUrl, "pml-training.csv")
  download.file(testUrl, "pml-testing.csv")
  
  # save download date
  dateDownloaded <- date()
}

# load the data with
training <- read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
testing <- read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))
```

The training data has a lot of NA's. We will exclude all variables with at least one NA, so the training of the model will be easier. 
The data also has a few variables that contain information about the test subject and the time when the data was gathered. We won't need these for our prediction. We will omit the first eight variables. 


```{r clean_data}
training2 <- training[, apply(training, 2, function(x) !any(is.na(x)))]
training3 <- training2[,-c(1:8)]
testing2 <- testing[,-c(1:8)]
dim(training3)
```


##Data partitioning and modelling

Next we will partition the training data into an training and validation set, so we can evaluate our model before applying it to the 20 test records.

```{r partition}
set.seed (1536)
inTrain <- createDataPartition(y=training3$classe, p = 0.75, list = FALSE)
trainSet <- training3[inTrain,]
validationSet <- training3[-inTrain,]

# Dimensions
dim(trainSet)
dim(validationSet)

```


Now we can build the model. We will use Random Forest with K-fold cross-validation with 5 folds.

```{r modelfit}
set.seed(402)
trainingControls <- trainControl(method = "cv", number = 5, allowParallel = TRUE, verbose = TRUE)
modfit <- train(classe ~ ., method="rf", data=trainSet, trControl = trainingControls, verbose = FALSE)
```

## Predictions

Let's make some predictions.

```{r prediction}
pred <- predict(modfit,validationSet)

confusionMatrix(pred, validationSet$classe)
```

We have achieved a model with 99.2% accuracy. 

